Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-05-09 src/tex/ms.tex
new references

% Previous studies
Previous studies have constrained amplitude birefringence by performing different statistical analyses.
\citet{Yamada_2020} and \citet{Wang_2021} both performed \ac{PE} on the events in the first \ac{GW} transient catalog \citep{GWTC-1}, GWTC-1, using birefringent templates.
\citet{Okounkova_2022} considered the distribution of observed inclinations of the \ac{GW} events in the second \ac{GW} transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence.

@@

@@

We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
This measurement is significantly more stringent than past constraints (Table \ref{tab:comparison_summary}).
As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources (Sec.~\ref{sec:results:hier}).
The result of that analysis was consistent with a vanishing $\kappa$ for all events in our catalog within 90\% credibility, but hinted at some possible variance in the population.
Future catalogs with more events will shed light on this feature.

>>>>

% Previous studies
Previous studies have constrained amplitude birefringence by performing different statistical analyses.
\citet{Yamada_2020} and \citet{Wang_2021} both performed \ac{PE} on the events in the first \ac{GW} transient catalog \citep{GWTC-1}, GWTC-1, using birefringent templates.
\citet{Okounkova_2022} considered the distribution of observed inclinations of the \ac{GW} events in the second \ac{GW} transient catalog \citep{GWTC-2}, GWTC-2, to look for signs of birefringence.
While this manuscript was finalized, \citet{Zhu:2023wci} reported an analysis of GWTC-3.

@@

The parameterization of \citet{Wang_2021} is the same as that in the more recent work by \citet{Zhu:2023wci}.
That work, which appeared while this manuscript was finalized, reported a constraint of $M_{\rm PV} > 4.1 \times 10^{-22}$ based on GWTC-3.

@@

We found no evidence of amplitude birefringence in the GWTC-3 data, and constrained \variable{output/restricted_kappa_median.txt}with 90\% credibility when treating $\kappa$ as a global quantity shared by all GWTC-3 events (Sec.~\ref{sec:results:gwtc}).
This measurement is significantly more stringent than past constraints (Table \ref{tab:comparison_summary}).
As an additional check, we implemented a hierarchical analysis that allowed for each event to probe different effective values of $\kappa$, as might be the case if birefringence is mediated by a field that is nonuniform over the angular separations between detected sources (Sec.~\ref{sec:results:hier}).
The result of that analysis was consistent with a vanishing $\kappa$ for all events in our catalog within 90\% credibility, but hinted at some possible variance in the population.
Future catalogs with more events will shed light on this feature, and enable richer models with explicit correlations between birefringent attenuation and source parameters like sky location \citep{Goyal:2023uvm,Ezquiaga:2021ler}.
